Supplementary Figures relative to filtering

### 2020-05-08 main_figures/dimensionality.Rmd
some figure updates

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
})
source("../misc_functions.R")
source(system.file("extdata/scrna_alternatives.R", package="pipeComp"))
theme_set(theme_cowplot(font_size = 12))
```

```{r, eval=FALSE}
# NB: the results were generated using an older version of pipeComp (branch 
# `biorxiv`), and must be plotted using this version of the package
suppressPackageStartupMessages({
  library(pipeComp)
})
```

@@

```{r, include=FALSE}
# the second part of these results was generated with pipeComp >0.99.26
devtools::load_all("../../pipeComp/")
```

```{r, fig.width=8, fig.height=4}
res <- readRDS("../data/results_dimensionality.rds")
H <- renameHrows(
  scrna_evalPlot_silh( res, step="dimensionality", reorder_rows=TRUE, 
                       row_split=gsub("norm\\.","",norm),
                       heatmap_legend_param=list(direction="horizontal")),
  function(x) gsub("norm\\.","",x))
p2 <- grid.grabExpr(draw(H, heatmap_legend_side="bottom"))

p3 <- renameHrows(
  evalHeatmap( res, what=c("NMI","min_pr"), agg.by=c("norm","dims"), 
                   row_split=gsub("norm\\.","",norm), reorder_rows=TRUE, anno_legend=FALSE) + 
  evalHeatmap(res, what=c("ARI"), agg.by=c("norm","dims"), row_split=gsub("norm\\.","",norm), 
              reorder_rows=TRUE, filterExpr=n_clus==true.nbClusts, 
              anno_legend=FALSE, title="ARI at\ntrue k"),
  function(x) gsub("norm\\.","",x)
)
p3 <- grid.grabExpr(draw(p3))
```

```{r figure6, fig.width=9, fig.height=9.5}
pdf("dimensionality.pdf", width=9, height=10)
plot_grid( p1, p2, p3, nrow=3, rel_heights=c(0.36,0.32,0.32), labels=c(NA,LETTERS[3:4]))
dev.off()
plot_grid( p1, p2, p3, nrow=3, rel_heights=c(0.36,0.32,0.32), labels=c(NA,LETTERS[3:4]))
```

>>>>

```{r}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(ComplexHeatmap)
  library(matrixStats)
  library(pipeComp) # the first part requires pipeComp >0.99.26
})
source("../misc_functions.R")
theme_set(theme_cowplot(font_size = 12))
```

```{r, fig.width=8, fig.height=4}
rnfn <- function(x){
  x <- gsub("norm\\.","",x)
  x <- gsub("maxLikGlobal$","maxLikGlobal(k=20)",x)
  gsub("maxLikGlobal10$","maxLikGlobal(k=10)",x)
}
res <- readRDS("../data/results_dimensionality.rds")
H <- renameHrows(
  scrna_evalPlot_silh( res, step="dimensionality", reorder_rows=TRUE, 
                       row_split=gsub("norm\\.","",norm),
                       heatmap_legend_param=list(direction="horizontal")),
  rnfn)
p2 <- grid.grabExpr(draw(H, heatmap_legend_side="bottom"))

p3 <- renameHrows(
  evalHeatmap( res, what=c("NMI","min_pr"), agg.by=c("norm","dims"), 
                   row_split=gsub("norm\\.","",norm), reorder_rows=TRUE, anno_legend=FALSE) + 
  evalHeatmap(res, what=c("ARI"), agg.by=c("norm","dims"), row_split=gsub("norm\\.","",norm), 
              reorder_rows=TRUE, filterExpr=n_clus==true.nbClusts, 
              anno_legend=FALSE, title="ARI at\ntrue k"),
  rnfn
)
p3 <- grid.grabExpr(draw(p3))
```

```{r, eval=FALSE}
# NB: the next results were generated using an older version of pipeComp (branch 
# `biorxiv`), and must be plotted using this version of the package
suppressPackageStartupMessages({
  library(pipeComp)
})
```

@@

```{r figure6, fig.width=9, fig.height=9.5}
pdf("dimensionality.pdf", width=9, height=10)
plot_grid( p1, p2, p3, nrow=3, labels=c(NA,LETTERS[3:4]))
dev.off()
plot_grid( p1, p2, p3, nrow=3, labels=c(NA,LETTERS[3:4]))
```

### 2020-05-08 supplementary_figures/supp2_filtering.Rmd
some figure updates

```{r, include=FALSE}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(SingleCellExperiment)
  library(reshape2)
  library(ComplexHeatmap)
  library(knitr)
  library(circlize)
  library(pipeComp)
})
source("../misc_functions.R")
```

@@

>>>>

```{r, include=FALSE}
suppressPackageStartupMessages({
  library(ggplot2)
  library(cowplot)
  library(SingleCellExperiment)
  library(reshape2)
  library(ComplexHeatmap)
  library(knitr)
  library(circlize)
  library(pipeComp)
  library(grid)
})
source("../misc_functions.R")
```

@@

```{r, fig.height=8.5, fig.width=7}
res <- readRDS("../data/filtering_intermediateSummary.rds")
tmp <- res$filtering$pcCells
dbl <- 1+grepl("doublet",row.names(tmp))
row.names(tmp) <- paste0(c("","doublet + ")[dbl],sapply(strsplit(row.names(tmp)," "), FUN=function(x) x[2]))
cols <- colorRamp2( seq(from=0,to=sqrt(max(tmp)),length.out=100)^2,
                    colorRampPalette(c("white","red"))(100) )
H <- Heatmap(tmp, bottom_annotation=pipeComp:::.ds_anno(colnames(tmp)), 
        show_column_names = FALSE, cluster_columns=FALSE, show_row_dend=FALSE, 
        row_split=c("No doublet removal","Doublet removal")[dbl], 
        name="proportion\nexcluded", col=cols, heatmap_legend_param=list(
          at=c(0,5,10,30,60), labels=paste0(c(0,5,10,30,60),"%"),
          grid_height=unit(7,"mm")) )

res <- readRDS("../data/overall.rds")
# get cell numbers
nb <- aggregate(res$evaluation$doublet$N.before, res$evaluation$doublet[,"dataset",drop=FALSE], FUN=sum)
row.names(nb) <- nb[,1]
res <- readRDS("../data/excluded.rds")
res <- lapply(res, FUN=function(x){
  names(x) <- gsub("doubletmethod=none;filt=filt.","",names(x))
  x <- x[names(x)!="doubletmethod=none;filt=none"]
  names(x) <- c("Doublet removal", "Default filter", "Stringent filter")
  x
})
res2 <- lapply(names(res), FUN=function(x){
  e <- sapply(res[[x]], FUN=function(i) sapply(res[[x]], FUN=function(j) sum(i %in% j)))
  e/nb[x,2]*100
})
names(res2) <- names(res)
h2 <- Heatmap(do.call(cbind, res2), column_split = rep(names(res2), each=ncol(res2[[1]])), 
              cluster_columns = FALSE, cluster_rows = FALSE, col=c("white","red"), 
              name="% of\ndataset", column_title_gp = gpar(font.size=10), column_title=" ",
              top_annotation=pipeComp:::.ds_anno(rep(names(res),each=ncol(res2[[1]])),legend = FALSE))
```

```{r, fig.height=6.5, fig.width=8}
plot_grid( grid.grabExpr(draw(H,merge_legend=TRUE)), grid.grabExpr(draw(h2)), 
           nrow=2, rel_heights=c(6,4), labels=LETTERS[1:2])
```

**A:** Proportion of cells filtered out by subpopulation. Applying the same filters in a cluster-wise fashion (using `scran::quickCluster`, and desginated here with `clustFilter*`) leads to virtually no cell exclusion. The color-mapping is square-root transformed to improve the visibility of differences at low proportions. **B:** Overlap between cells excluded by doublet removal (`scDblFinder`) and those excluded by MAD-based filters (without doublet removal; the filters are described in the methods), expressed as a proportion of the dataset. The cells excluded as doublets do not tend to be excluded by (even stringent) MAD-based filtering.
